---
title: "p8105_hw6_xw3038"
author: "Avery Wang"
date: "2024-12-01"
output: github_document
---


Load the packages

```{r}
library(tidyverse)
library(broom)
library(purrr)
library(modelr)
library(rnoaa)
set.seed(123)
```

```{r}
library(p8105.datasets)
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


## Problem 2

* Load and Clean the data

```{r}
homicide=read_csv(file = "./data/homicide-data.csv", 
                  na = c(".", "NA", "Unknown")) |>
  janitor::clean_names()

homicide_cleaned =homicide |>
  mutate(
    city_state=paste(city, state, sep = ", "),
    is_solved=if_else(disposition=="Closed by arrest", 1, 0),
    victim_age=as.numeric(victim_age),
    victim_race=as.factor(victim_race),
    victim_sex=as.factor(victim_sex)
  ) |>
  # Exclude specific cities and limit analysis to specified victim races
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", 
                        "Tulsa, AL")) &
      victim_race %in% c("White", "Black"),
    !is.na(victim_sex), !is.na(victim_age), !is.na(victim_race), 
    !is.na(is_solved)
  )

```

* Extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims

```{r}
baltimore_data=homicide_cleaned |>
  filter(city_state=="Baltimore, MD")
glm_model=glm(
  is_solved ~ victim_age + victim_sex + victim_race,
  family = binomial(link = "logit"),
  data = baltimore_data
)

glm_summary_odds=broom::tidy(glm_model,conf.int = TRUE, exponentiate = TRUE)
glm_summary_odds

```

```{r}
male_adjusted_odds=glm_summary_odds|>
  filter(term == "victim_sexMale")|>
  select(term, estimate, conf.low, conf.high)
# View the result
print(male_adjusted_odds)
```

* Run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. 

```{r}
city_glm=homicide_cleaned|>
  group_by(city_state)|>
  nest()|>
  #add model for each city
  mutate(glm_model=map(data, \(df) glm(is_solved ~ victim_age + 
                                         victim_sex + victim_race, 
                                family = binomial(link = "logit"), 
                                data = df)),
         result=map(glm_model, \(model) broom::tidy(model, 
                                                    conf.int = TRUE, 
                                                    exponentiate = TRUE)))|>
  select(-data,-glm_model)|>
  unnest(result)
```

```{r}
male_city_glm=city_glm|>
  filter(term == "victim_sexMale") |>
  select(city_state, estimate, conf.low, conf.high)
male_city_glm
```

* Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
city_results_arrange=male_city_glm|>
  arrange(estimate)|>
  mutate(city_state = factor(city_state, levels = city_state))
city_results_arrange
```

```{r}
city_results_arrange|>ggplot(aes(x=reorder(city_state, estimate), y = estimate, 
                                 color=city_state)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  coord_flip() + 
  labs(
    title = "Adjusted Odds Ratios (Male vs Female) by Different City",
    x = "City",
    y = "Odds Ratio"
  ) +
  theme_minimal()+ 
  theme(legend.position = "none")
```

Cities with OR > 1 (such as Stockton, CA; Albuquerque, NM) show the cases involving male victims are more likely to be solved than those involving female victims. And the majority cities have OR less than 1. Some cities, like Stockton, CA and Albuquerque, NM, have wide CIs, indicating high variability and uncertainty. Narrower CIs in cities like Chicago, IL also suggest more reliable estimates.

## Problem 3

```{r}
birthweight=read_csv(file = "./data/birthweight.csv", 
                     na = c(".", "NA", "Unknown")) |>
  janitor::clean_names()
```

* Load and clean the data for regression analysis

```{r}
birthweight_cleaned=birthweight|>
  mutate( 
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), 
                   labels = c("White", "Black", "Asian", 
                              "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", 
                              "Puerto Rican", "Other")),
    malform = factor(malform, levels = c(0, 1), 
                     labels = c("Absent", "Present")),
    fincome = fincome * 100
  )|>
  drop_na()
```

* Propose a regression model for birthweight

```{r}
birthweight_model=lm(
  bwt ~ babysex + blength + bhead + momage + mheight + ppbmi +
    smoken + wtgain + menarche+ fincome,
  data = birthweight_cleaned
)
birthweight_model
```

The goal is to predic the baby's birthweight (`bwt`) assuming that the birthweight is influenced by:
Baby's Physical Characteristics: `babysex`, `blength`, `bhead`. Maternal Characteristics: `momage`, `mheight`, `ppbmi`, `smoken`, `wtgain`, `menarche`. Family condition as the sociaeconomic consideration: `fincome`. 

```{r}
birthweight_cleaned |>
  add_predictions(birthweight_model) |>
  add_residuals(birthweight_model)|>
  ggplot( aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Predictions ",
    y = "Residuals"
  ) +
  theme_minimal()
```


* Compare with the other two models


```{r}
model2=lm(bwt ~blength + gaweeks, data = birthweight_cleaned)
model3=lm(bwt~ bhead * blength * babysex, data = birthweight_cleaned)

```

```{r}
cv_df = crossv_mc(birthweight_cleaned, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
)

cv_df=cv_df |> 
  mutate(
    # Define and fit models
    original_mod = map(train, \(df) lm(bwt ~babysex + blength + bhead + 
                  momage + mheight + ppbmi +
                  smoken + wtgain + menarche+ fincome, 
                  data = df)),
    model2_mod = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model3_mod = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))
  )|>
  mutate(
    # Calculate RMSE for each model on the test sets
    rmse_original = map2_dbl(original_mod, test, \(mod, df) rmse(model = mod, 
                                                                 data = df)),
    rmse_model2 = map2_dbl(model2_mod, test, \(mod, df) rmse(model = mod, 
                                                             data = df)),
    rmse_model3 = map2_dbl(model3_mod, test, \(mod, df) rmse(model = mod, 
                                                             data = df))
  )

```

```{r}
summary_results=cv_df |> 
  summarise(
    mean_rmse_original = mean(rmse_original),
    mean_rmse_model2 = mean(rmse_model2),
    mean_rmse_model3 = mean(rmse_model3)
  )

print(summary_results)
```
```{r}
cv_long=cv_df |> select(starts_with("rmse"))|>
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model))


```

```{r}
 cv_long|>ggplot(aes(x = model, y = rmse)) + 
   geom_violin()+
  labs(
    title = "Cross-Validated RMSE for Birthweight Models",
    x = "Model",
    y = "RMSE"
  ) +
  theme_minimal()
```

The original hypothesized model seems to have the best performance, with relatively the best accuracy. The model with the main effect shows the highest rmse, indicating it performs worse. And the model three-way interaction performs slightly better but still wors than the orginial hypothesized model